---
title: "Week 5 Assignment"
author: "Dan Brooks"
date: "February 28, 2016"
output: html_document
---
#Extract the CSV file from my Github Page and get all of the libraries that are needed for this assignment.
```{r}
library(RCurl)
library(dplyr)
library(tidyr)
URL <- getURL("https://raw.githubusercontent.com/DanielBrooks39/IS607/master/Week_5/Flight_Information.csv")
FlightData <- read.csv(text = URL, header = TRUE)
```
***
#Gives names to the vectors that are in the data frame and add information to the table where there is a blank spot.
#Create a Tidy dataset with the columns Airline, Info, Destinations and the total flights that were delayed or ontime for waech airline and destination
```{r}
names(FlightData) <- c("Airline", "Info", "Los Angeles","Phoenix", "San Diego", "San Francisco", "Seattle")
FlightData$Airline[2] <- "Alaska"
FlightData$Airline[4] <- "AM West"
Tidy <- FlightData %>% gather("Destination", "Timing", 3:7)
```
***
#Separate the full tidy dataset into delayed and ontime flights 
```{r}
Delay <- Tidy %>% filter(Info == "Delay")
OnTime <- Tidy %>% filter(Info == "OnTime"|Info == "Ontime")
```
***
#Find the mean number of lfights per airline that is ontime and delayed
```{r}
AvgDelay <- Delay %>% group_by(Airline) %>% summarise(mean = mean(Timing))
AvgDelay
AvgOntime <- OnTime %>%  group_by(Airline) %>% summarise(mean = mean(Timing))
AvgOntime
```
***
#Join the delayed and ontime datasets together and figure out the ration between the number of ontime flights to the number of delayed flights per airline
```{r}
Joined <- inner_join(AvgDelay, AvgOntime, by = "Airline")
names(Joined) <- c("Airline", "AvgDelay", "AvgOnTime") 
FlightInfo <- Joined %>% mutate("Ratio(OnTime/Delay)" = AvgOnTime/AvgDelay)
FlightInfo
```
***
#Find the mean number of flights that were ontime or delayed according to their destination
```{r}
AvgDelay <- Delay %>% group_by(Destination) %>%  summarise(mean = mean(Timing))
AvgDelay
AvgOntime <- OnTime %>%  group_by(Destination) %>% summarise(mean = mean(Timing))
AvgOntime
```
***
#Find the ratio between the avg number of flights that were ontime and the average number of flights that were delayed based on their destination
```{r}
Joined <- inner_join(AvgDelay, AvgOntime, by = "Destination")
names(Joined) <- c("Destination", "AvgDelay", "AvgOnTime")
DestInfo <- Joined %>% mutate("Ratio" = AvgOnTime/AvgDelay) %>% arrange(desc(Ratio))
DestInfo
```

#If I had to pick an airline to fly on, I would pick AM West, because their ratio of ontime flights to delayed flights is higher thatn Alaska. Also, I would pick Phoenix as my destination because their ratio is also the best compared to the other destinations.
