---
title: "Week 4 Assignment"
author: "Dan Brooks"
date: "February 14, 2016"
output: pdf_document
---
This may be a little bit of over kill, but I wanted to try and extract the first
names and then last names separately to practice my extracting skill. I then 
concatenated the two strings together to get the first name last name order
that the assignment calls for.

***
```{r name_vector}
library(stringr)
name <- c("Moe Szyzlak", "Burns, C. Montgomery", "Rev. Timothy Lovejoy", "Ned Flanders", "Simpson, Homer", "Dr. Julius Hibbert")
name
```

```{r question_1a}
first_name <- unlist(str_extract_all(toupper(name), 
"M..[[:space:]]|R..\\.[[:space:]][[:alpha:]]{7}|NED|D.\\.[[:space:]][[:alpha:]]{6}|C[[:punct:]][[:space:]][[:alpha:]]{10}|H[E-R]{4}"))
first_name
```

```{r last_name}
last_name <- unlist(str_extract_all(toupper(name), 
"SZ[[:alpha:]]{5}|B[N-U]{4}|LO.+|F[[:alpha:]]{7}|SI.{5}|HI.{5}"))
last_name
```

```{r string_concatenation}
full_name <- str_c(first_name, last_name, sep = " ")
full_name
```

***

``` {r question_1b}
char_vector <- str_detect(toupper(full_name), "[[:alpha:]]{2,3}[[:punct:]]")
char_vector
```

```{r question_1c}
##I am assuming that a second name is anybody that has more than two parts to their name.
##Homer Simpson only has one name, but Dr. Julius Hibbert has more than one.

second_name <- str_detect(toupper(full_name), "[[:alpha:]]{1,3}[[:punct:]][[:space:]][[:alpha:]]")
second_name
```

***
```{r question_2}
vector <- c("<title>+++BREAKING NEWS+++</title>")
HTMLtag <- unlist(str_extract_all(vector, "<.+>"))
HTMLtag
##This fails because the (.) means that it can be any character. The (+) sign
##after that means that the wildcard can be matched one or more times. That 
##means that and character can be macthed any amount of times. That will give
##you the entire enxpression. The <> characters really do not matter in this case.
```
```{r FIX}
HTMLtagfix <- unlist(str_extract_all(vector,"<[[:alpha:]]+>"))
HTMLtagfix
```

***

```{r question_3}
binomial_theorem <- c("(5-3)^2=5^2-2*5*3+3^2")
extract <- unlist(str_extract_all(binomial_theorem, "[^0-9=+*()]+"))
extract
##This fails because there is no \\ before the special characters.
##^ and (-) need to be \\^ and \\- for the expression to work.
```
```{r FIX_2}
extract_expression <- unlist(str_extract_all(binomial_theorem, "[\\^\\-[0-9]=+*()]"))
extract_expression
```

***

```{r question_4}
Code <- c('clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0TanwoUwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfex.r1w1YwwojigOd6vrfUrbz.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr')
numbers <- unlist(str_extract_all(Code, "[0-9]"))
numbers
lower_case <- unlist(str_extract_all(Code, "[a-z]"))
lower_case
upper_case <- unlist(str_extract_all(Code, "[A-Z\\!]"))
upper_case
```
I decided to start off by parsing out the code by upper case, lower case and
numbers. I figured that would be a good way to start. It turns out that the
answer just happens to be in all captial letters. Makes the answer pretty
much stand right out.
The answer is:
CONGRATULATIONS YOU ARE A SUPER NERD!
